---
title: "Employee motivation translation scheme"
description: |
  The full SEM2Agent scheme for the employee motivation model.
author:
  - name: Laura Burbach
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(distill)
library(kableExtra)
options(knitr.table.format = "html")
```

This is an application of the SEM2Agent scheme for an ABM based on the SEM in:

Rožman, M., Tominc, P. & Milfelner, B. (2020). A Comparative Study Using Two SEM Techniques on Different Samples Sizes for Determining Factors of Older Employee’s Motivation and Satisfaction. Sustainability 2020, 12(6), 2189; https://doi.org/10.3390/su12062189

# Overview
## Purpose and patterns
The purpose of the model is twofold. 
First, it considers how the stress perception of an agent and the stress departing from a company as well as the perception of the relations at the workplace and of how nice other employees are (Niceness Fraction) influence how satisfied and motivated (elder) employees are. It also considers how the satisfaction of employees influences the motivation of employees. 
Where employees are stressed, negative employee relationships arise and employees become less nice. Additionally, higher levels of stress also lead to lower employee satisfaction. Good relations in the workplace lead to an increased level of satisfaction. Favorable and positive relations also lead to a higher level of work motivation. In addition, there is a positive connection between employee satisfaction and motivation.
Second, the model considers, whether the company success is higher or lower when a company values their employees and managers guard their employees against high amounts of company stress. The model shows, whether it is better to increase the company stress to reach higher success or whether company stress must be reduced to keep the motivation of employees high and thus reach more success. 

A possible pattern to evaluate this against would be empirical data on employees stress, relations, work motivation, satisfaction and company success. 

## Entities, state variables and scales
The model includes two entities, the employees (huamn agents, dynamic), the company (workplace, dynamic). 

The following table shows all entities with their state variables:

```{r}
# You can prepare the table as a csv and upload it here.
read.csv2("tables/StateVariablesEmployees.csv") %>%
  as.data.frame() %>%
  kbl(col.names = c(
            "Entity",
            "State variable",
            "Represents",
            "Dynamic / static",
            "Type",
            "Range / value"
          ))  %>%
  kable_styling()
```						

Representation of time:
Each model tic represents one work week. 
The model can run for a longer time, but we conducted experiments with 50 steps (50 weeks).

Representation of space:
Space is represented in a simplified way.
The agents are placed on a barabasi albert grid so that each agent has eight neighbors.
The agents do not move on the grid.

## Process overview and scheduling
In each model step, all agents are activated in random order.
1. Agents are activated. The agent modifies his/her niceness: if the stress is very high, they get not nice. If the stress is very low, they get nice. Otherwise, they adjust their niceness to the niceness of a random neighbor. 
2. The agents adjust their employee relations based on the agent stress and the fraction of nice neighbors.
3. The agents adjust their employee satisfaction based on the agent stress and employee relations.
4. The agents adjust their employee motivation based on the employee relations and employee satisfaction. 
5. The agents adjust their agent stress based on their previous agent stress combined with the company stress. 
6. The company stress is updated based on the company success and the degree of employee valuation. 
7. The company success is updated based on the employee motivation. 

# Design concepts
## Basic principles
The model is based on the SEM by Rožman, Tominc and Milfelner (2020). Based on this structural equation model, it models the influence of stress on relations and satisfaction, the influence of relations on satisfaction and motivation and how motivation and satisfaction influence each other.

Additionally, the model considers how the niceness of employees influences the motivation of employees.

Further, it considers how it influences the employee motivation and the company success when companies value their employees (low, medium, high) and managers guard their employees from stress or do not. 

## Emergence
The emergent system behavior is how stongly the perceptions of relations and of stress influence the satisfaction and in turn how strongly satisfaction influences motivation and how stronly stress influences relations and how strongly niceness influences motivation. 
Further, the emergent system behavior is how strongly the valuation of employees by reducing stress or not influences company success and employee motivation.

Overall, the emergence is that we can observe which variable(s) has/have the strongest influence and when agents are the most satisfied and motivated as well as when the company success is the highest/lowest. 

## Individual decision making
Agents make no intentionally decisions. 

## Adaptation
There is no direct adaptation. There is no direct objective-seeking.

## Objectives
The agents do not have objectives other than being motivated and satisfied, but no intentional objectives.

## Learning
Learning is not included.

## Prediction
Prediction is not explicitly included. Implicitly, the agents predict that changing levels of stress and relations change their satisfaction.

## Sensing
Agents are assumed to exactly sense all external and internal variables. However, that variables input on their state variables is modulated by path coefficients.

## Interaction
In each week, an agent interacts profoundly with one other agent in our model. Additionally, they perceive the overall employee relations.  

## Heterogeneity
Agents are homogenous in their decision-making. 

## Stochasticity
Stochasticity is included to generate variability in the initial value of the agent‘s state variables. The values are drawn from a distribution based on the mean and standard deviation values reported in the model (means must be calculated using the single item means). This is done to represent different work-situations (with more/less stress, better/worse relations, different conditions that influence satisfaction). 

## Collectives
There are no collectives in the model.

## Observation
For each model step, we track the stress, niceness, relations, satisfaction and motivation of the agents. Likewise, we track the company stress and the company success.

# Details
## Implementation details
The model is implemented in Julia using the Agents.jl package. It is accessible (on GitHub).

## Initialization
Upon initialization, the following entities are created:
* The employee agent.
* The company (workplace).

The employee agent is created with the following state variables: 
* Stress, sourced from empirical data.
* Relations, sourced from empirical data.
* Satisfaction, sourced from empirical data.
* Motivation, sourced from empirical data.
* NicenessFraction: 0...1.


The company is created with the following state variables:
* Success: -1...1.
* Employee valuation: low, medium, high. 

The employee agents with all values for all state variables drawn from distributions based on the mean survey results reported in the paper. This variability is supposed to represent variability in workplaces or in the perception of employees.

## Input Data
### Overview
The data for the agent‘s state variables comes from the paper by Rožman, Tominc and Milfelner (2020). Available data is only the means and sds from the single items of the four variables and the path coefficients.

### Structure
In the paper, there are three main types of data: Descriptive survey data for all variables in the model (Stress, Relations, Satisfaction, Motivation), path coefficients for the relationships between the variables. CB-SEM:


```{r}
# You can prepare the table as a csv and upload it here.
read.csv2("tables/InputDataStructure.csv") %>%
  as.data.frame() %>%
  lapply(function(x) {gsub("\\*", "\\\\*", x)}) %>%
  as.data.frame() %>%
  kbl(col.names = c(
            "Variable",
            "SEM Data",
            "Range"
          ))  %>%
  kable_styling()
```				

### Mapping
All SEM variables are selected as state variables characterizing the agent‘s motivation. 
The agents are initialized with a value which is randomly sampled based on the descriptive statistics.
The structural equations with the associated path coefficients are used for each agent to compute the new values for the variables in each time step.
The initial values are normalized so all scales reach from -1 to 1.
Similarly, most path coefficients are normalized so their sum is 1.


```{r}
# You can prepare the table as a csv and upload it here.
read.csv2("tables/mappingEmployeeMotivation.csv") %>%
  as.data.frame() %>%
  lapply(function(x) {gsub("\\*", "\\\\*", x)}) %>%
  as.data.frame() %>%
  kbl(col.names = c(
            "Source variable",
            "ABM state variable",
            "Range",
            "Initialization",
            "Updating"
          ))  %>%
  kable_styling()
```



### Patterns
The regression equations from the SEM are used to calculate the model variables in each step.


## Submodels
There are no submodels.
