---
title: "Yolo farming ABM results"
description: |
  The results of the climate program in farming model.
author:
  - name: Lilian Kojan
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(distill)
library(knitr)
library(dplyr)
library(DT)
library(arrow)
library(here)
```

These are the results of the application of the Yolo farming agent-based model.
This model serves as a demonstration of SEM2Agent scheme for an ABM based on the SEM in:

Niles, M. T., Lubell, M., & Haden, V. R. (2013). Perceptions and responses to climate policy risks among california farmers. Global Environmental Change, 23(6), 1752–1760. https://doi.org/10.1016/j.gloenvcha.2013.08.005

# Model parameters
Each model is initialized with 100 agents and replicated 500 times.
The results presented here are averaged both over all agents, as well as all replications.
In the experiments, three model parameters can be manipulated:
The threshold of neighbors needed to exert normative pressure to participate (default: 0.5), the approximate fraction of agents originally participating in the program (default: 0.25) and the overall program quality (default: 0.4).

# Experiments
## Experiment 1

The first experiment runs with the default parameters.

```{r}
# Upload results table.
read_arrow(here::here("model_yolo-farming", "data-output", "experiment-1_summary.arrow")) %>%
  as.data.frame() %>%
  mutate(across(c(2:18), ~formatC(round(.x, 2), 2, format = "f", replace.zero = TRUE))) %>%
  transmute(Step = step,
            Year = year,
            `Water availability` = wateravailability,
            `Climate change experience` = paste0(ccexperience_mean, " (±", ccexperience_std, ")"),
            `Policy experience` = paste0(ppexperience_mean, " (±", ppexperience_std, ")"),
            `Climate change belief` = paste0(ccbelief_mean, " (±", ccbelief_std, ")"),
            `Perceived climate change risk` = paste0(ccrisk_mean, " (±", ccrisk_std, ")"),
            `Participation intention` = paste0(ppintention_mean, " (±", ppintention_std, ")"),
            `Fraction participating` = paste0(participation_mean, " (±", participation_std, ")"))  %>% 
  datatable(rownames = FALSE, caption = 'Experiment 1 summary results: means (± standard deviation) for 100 agents and 500 iterations', options = list(pageLength = 11))
```

```{r, out.width="l-body"}
# Upload figure
include_graphics(here::here("model_yolo-farming", "data-output", "experiment-1_figure.svg"))
```

As expected, the climate change experience curve runs in reverse to the exogenous water availability.
The policy experience weakly improves over time.
The fraction of participants takes a dip in the first step and then also grows.
On average, 68 % of farmer agents participate after 10 years.
The variance of participation is large and constant throughout all steps.
This might be due to the variation in initial conditions.

## Experiment 2

In the second experiment,the social threshold parameter is increased to 1.
This implies that all 4 neighbors of an agent have to participate for the neighbor to feel normative pressure to participate themselves.

```{r}
# Upload results table.
read_arrow(here::here("model_yolo-farming", "data-output", "experiment-2_summary.arrow")) %>%
  as.data.frame() %>%
  mutate(across(c(2:18), ~formatC(round(.x, 2), 2, format = "f", replace.zero = TRUE))) %>%
  transmute(Step = step,
            Year = year,
            `Water availability` = wateravailability,
            `Climate change experience` = paste0(ccexperience_mean, " (±", ccexperience_std, ")"),
            `Policy experience` = paste0(ppexperience_mean, " (±", ppexperience_std, ")"),
            `Climate change belief` = paste0(ccbelief_mean, " (±", ccbelief_std, ")"),
            `Perceived climate change risk` = paste0(ccrisk_mean, " (±", ccrisk_std, ")"),
            `Participation intention` = paste0(ppintention_mean, " (±", ppintention_std, ")"),
            `Fraction participating` = paste0(participation_mean, " (±", participation_std, ")"))  %>% 
   datatable(rownames = FALSE, caption = 'Experiment 2 summary results: means (± standard deviation) for 100 agents and 500 iterations', options = list(pageLength = 11))
```
```{r, out.width="l-body"}
# Upload figure
include_graphics(here::here("model_yolo-farming", "data-output", "experiment-2_figure.svg"))
```
The climate change experience curve remains broadly unchanged compared to default conditions, and the policy experience curve is also similar.
The participating fraction, on the other hand, declines in the firs step and then remains constant and low.
After 10 years, the participation fraction is on average 0.13.

## Experiment 3

In the third experiment, the fraction of initial participants is increased to 0.5.

```{r}
# Upload results table.
read_arrow(here::here("model_yolo-farming", "data-output", "experiment-3_summary.arrow")) %>%
  as.data.frame() %>%
  mutate(across(c(2:18), ~formatC(round(.x, 2), 2, format = "f", replace.zero = TRUE))) %>%
  transmute(Step = step,
            Year = year,
            `Water availability` = wateravailability,
            `Climate change experience` = paste0(ccexperience_mean, " (±", ccexperience_std, ")"),
            `Policy experience` = paste0(ppexperience_mean, " (±", ppexperience_std, ")"),
            `Climate change belief` = paste0(ccbelief_mean, " (±", ccbelief_std, ")"),
            `Perceived climate change risk` = paste0(ccrisk_mean, " (±", ccrisk_std, ")"),
            `Participation intention` = paste0(ppintention_mean, " (±", ppintention_std, ")"),
            `Fraction participating` = paste0(participation_mean, " (±", participation_std, ")"))  %>% 
  datatable(rownames = FALSE, caption = 'Experiment 3 summary results: means (± standard deviation) for 100 agents and 500 iterations', options = list(pageLength = 11))
```
```{r, out.width="l-body"}
# Upload figure
include_graphics(here::here("model_yolo-farming", "data-output", "experiment-3_figure.svg"))
```
Compared to experiments 1 and 2, the fraction of participants increases quickly and strongly until, after 10 years, most agents (M = 0.82) participate in the program.
The slope of the policy experience curve is slightly less steep, and the variance in experiences is less pronounced.
This corresponds to the agents having more similar policy experiences, as a larger fraction of them participats in the same new program.

## Experiment 4

For the last experiment, we set the program quality parameter to -0.4 to reflect a program with overall low quality, e.g., not being efficacious in combating climate, or being overly costly for farmers.

```{r}
# Upload results table.
read_arrow(here::here("model_yolo-farming", "data-output", "experiment-4_summary.arrow")) %>%
  as.data.frame() %>%
  mutate(across(c(2:18), ~formatC(round(.x, 2), 2, format = "f", replace.zero = TRUE))) %>%
  transmute(Step = step,
            Year = year,
            `Water availability` = wateravailability,
            `Climate change experience` = paste0(ccexperience_mean, " (±", ccexperience_std, ")"),
            `Policy experience` = paste0(ppexperience_mean, " (±", ppexperience_std, ")"),
            `Climate change belief` = paste0(ccbelief_mean, " (±", ccbelief_std, ")"),
            `Perceived climate change risk` = paste0(ccrisk_mean, " (±", ccrisk_std, ")"),
            `Participation intention` = paste0(ppintention_mean, " (±", ppintention_std, ")"),
            `Fraction participating` = paste0(participation_mean, " (±", participation_std, ")"))  %>% 
  datatable(rownames = FALSE, caption = 'Experiment 4 summary results: means (± standard deviation) for 100 agents and 500 iterations', options = list(pageLength = 11))
```
```{r, out.width="l-body"}
# Upload figure
include_graphics(here::here("model_yolo-farming", "data-output", "experiment-4_figure.svg"))
```
For this specification both the participating fraction and the policy experience remain low and constant.
On average, after 10 years 21 % of participants participate, which is a decline from the initial 25 %.


# Summary
```{r}
include_graphics(here::here("model_yolo-farming", "data-output", "experiments-summary_figure.svg"))
```

The experiments re-emphasize the results already discussed by Niles et al., namely that the climate change experience is only one of many factors influencing farmers' participation in climate protective government programs.
As demonstrated in experiment 4, a program that is perceived to be ineffective or unfair will engender low participation, even if farmers feel the effects of climate change.
We added subjective norm as a driver for participating intention to complement the path from (perceived) climate change risk to government program participation (intention).
Varying the influence of subjective norm (experiment 2 and 3) results in strong changes in overall participation.
While subjective norm can be seen as a placeholder for other drivers of participation, this re-emphasizes the importance of examining what those drivers are and whether they can be utilized to increase government program participation.