---
title: " translation scheme"
description: |
  The full SEM2Agent scheme for .
author:
  - name:
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(distill)
library(kableExtra)
options(knitr.table.format = "html")
```

This is an application of the SEM2Agent scheme for an ABM based on the SEM in:

# Overview
## Purpose and patterns



## Entities, state variables and scales


The following table shows all entities with their state variables.

```{r}
# You can prepare the table as a csv and upload it here.
read.csv2() %>%
  as.data.frame() %>%
  kbl(col.names = c(
            "Entity",
            "State variable",
            "Represents",
            "Dynamic / static",
            "Type",
            "Range / value"
          ))  %>%
  kable_styling()
```						

Representation of time:

Representation of space:



## Process overview and scheduling


# Design concepts
## Basic principles


## Emergence


## Individual decision making


## Adaptation


## Objectives


## Learning


## Prediction



## Sensing


## Interaction


## Heterogeneity


## Stochasticity


## Collectives


## Observation


# Details
## Implementation details


## Initialization


## Input Data
### Overview


### Structure


```{r}
# You can prepare the table as a csv and upload it here.
read.csv2() %>%
  as.data.frame() %>%
  lapply(function(x) {gsub("\\*", "\\\\*", x)}) %>%
  as.data.frame() %>%
  kbl(col.names = c(
            "Variable",
            "Range",
            "Descriptive statistics",
            "SEM regression equation",
            "SEM R^2"
          ))  %>%
  kable_styling()
```				

### Mapping


```{r}
# You can prepare the table as a csv and upload it here.
read.csv2() %>%
  as.data.frame() %>%
  lapply(function(x) {gsub("\\*", "\\\\*", x)}) %>%
  as.data.frame() %>%
  kbl(col.names = c(
            "Source variable",
            "ABM state variable",
            "Range",
            "Initialization",
            "Updating"
          ))  %>%
  kable_styling()
```



### Patterns


## Submodels
